---
title: "R Notebook"
author: "Shivani Sanjay Mahaddalkar"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---
##Libraries
Importing appropriate libraries
```{r}
require(caret)
require(e1071)
require(rpart)
require(dplyr)
require(stringr)
require(randomForest)
library(tidyverse)    
library(kernlab)      
library(e1071)        
library(RColorBrewer)
#library(rsample)
```


##Data
The analysis aims to build models to recognize the handwritten digits from 0-9. The data set comes from the Kaggle Digit Recognizer competition. We build KNN, SVM and random forests algorithms and compare their accuracies.
Setting directory from which to extract data
```{r}
setwd("D:/Spring2021/DataAnalytics")
digit_train <- read.csv("digit-train.csv")
digit_test <- read.csv("digit-test.csv")

```

##Data Transformation:
The label column which is the prediction column or the target column is changed from an integer type to a factor type to make the model understand that classification with 10 levels from 0-9 is taking place. 
The columns which show zero variance are dropped, as they will not impact the prediction.
```{r}
var0 <- apply(digit_train, 2, var)!=0
digit_train <- digit_train[,var0]
digit_test <- digit_test[,colnames(digit_train)]
digit_train$label <- factor(paste0('X', digit_train$label), levels = c('X0', 'X1', 'X2', 'X3', 'X4', 'X5', 'X6', 'X7', 'X8', 'X9'))
digit_test$label <- factor(paste0('X', digit_test$label), levels = c('X0', 'X1', 'X2', 'X3', 'X4', 'X5', 'X6', 'X7', 'X8', 'X9'))
```

## Model: KNN algorithm
```{r}
search_grid = expand.grid(k = c(5, 7, 9, 11))

# set up 3-fold cross validation procedure
train_control <- trainControl(
  method = "cv", 
  number = 3
  )

# more advanced option, run 3 fold cross validation 5 times
train_control_adv <- trainControl(
  method = "repeatedcv", 
  number = 3,
  repeats = 5
  )

# train model
knn <- train(label ~ .,
  data = digit_train,
  method = "knn",
  trControl = train_control_adv,
  tuneGrid = search_grid
  )
```

```{r}
# top model

knn$results %>% 
  top_n(1, wt = Accuracy) %>%
  arrange(desc(Accuracy))

# results for best model
confusionMatrix(knn)
#digit_train[,'pixel138']
pred <- predict(knn, newdata = digit_test)
```
```{r}
confusionMatrix(pred, digit_test$label)
```
##Model: SVM

```{r}
set.seed(123)
svm.m1 <- svm(label ~ ., data = digit_train, type = 'C', kernel = 'linear', cross = 3, probability = TRUE)
summary(svm.m1)
set.seed(123)
svm.m2 <- svm(label ~ ., data = digit_train, type = 'C', kernel = 'radial', cross = 3, probability = TRUE)
summary(svm.m2)

pred1 <- predict(svm.m1, newdata = digit_test)
pred2 <- predict(svm.m2, newdata = digit_test)
```
```{r}
confusionMatrix(pred1, digit_test$label)
confusionMatrix(pred2, digit_test$label)
```

```{r}
set.seed(123)
train_control <- trainControl(method = 'repeatedcv', number = 3, repeats = 3)
rf <- train(label ~ ., data = digit_train, method = 'rf', metric = 'Accuracy', trControl = train_control, type = 'C')
confusionMatrix(rf)

```

```{r}
pred3 <- predict(rf, newdata = digit_test)
confusionMatrix(pred3, digit_test$label)
```
